---
title: "Geog533 Lab 10"
author: "Michael Kezner"
date: "November 28, 2017"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

## <span style="color:blue">Question 1</span>
Use R package UScensus2010county to complete the following tasks:  (20 pt.)
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
if(!require(UScensus2010)) {install.packages("UScensus2010")}
library(UScensus2010)
if(!require(UScensus2010county)) {install.county("osx")}
library(UScensus2010county)
library(tmap)
```


### <span style="color:red">Question 1(a)</span>
Plot a map of New York counties using the plot function.
```{r}
data("new_york.county10")
shp <- new_york.county10
df <- shp@data
plot(shp)
```

### <span style="color:red">Question 1(b)</span>
Plot a map of New York counties using the qtm function.
```{r}
qtm(shp)
```


### <span style="color:red">Question 1(c)</span>	
How many counties in New York State?
```{r}
NROW(df)
```

### <span style="color:red">Question 1(d)</span>
What’s the 3-digit fips code of Broome County?
```{r}
df[1,4]
```

### <span style="color:red">Question 1(e)</span>
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
pop1 <- df$P0010001
sum(pop1)
min(pop1)
max(pop1)
mean(pop1)
median(pop1)
library(moments)
skewness(pop1)
```

### <span style="color:red">Question 1(f)</span>	
Create a histogram and a boxplot of the population.
```{r}
hist(pop1, main = "P0010001")
boxplot(pop1)
```


## <span style="color:blue">Question 2</span>
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r message=FALSE, warning=FALSE}
if(!require(UScensus2010tract)) {install.tract("osx")}
library(UScensus2010tract)
```

### <span style="color:green">Question 2(a)</span>	
Plot a map of New York census tracts using the plot function.
```{r}
data("new_york.tract10")
shp2 <- new_york.tract10
df2 <- shp2@data
plot(shp2)
```

### <span style="color:green">Question 2(b)</span>
Compute the total population based on census tracts.
```{r}
sum(df2$P0010001)
```

### <span style="color:green">Question 2(c)</span>
Select all census tracts in Broome County and plot the map. 
```{r}
broome <- df2$county == "007"
plot(shp2[broome,])
```

### <span style="color:green">Question 2(d)</span>
What’s the total population of Broome County?
```{r}
broomepop <- df2[df2$county == "007",]
sum(broomepop$P0010001)
```

### <span style="color:green">Question 2(e)</span>
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(broomepop$P0010001)
boxplot(broomepop$P0010001)
```

### <span style="color:green">Question 2(f)</span>
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r message=FALSE, warning=FALSE}
sel <- data.frame(broomepop[,1:5])
sel$ratio <- broomepop$P0010001/200600
sel
```


## <span style="color:blue">Question 3</span>

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r, message=FALSE, warning=FALSE}
library(raster)
library(ncdf4)
```

### <span style="color:purple">Question 3(a)</span>		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
ndvi <- raster("NDVI.nc")
```

### <span style="color:purple">Question 3(b)</span>	
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(ndvi)
ncol(ndvi)
ncell(ndvi)
nbands(ndvi)
res(ndvi)
extent(ndvi)
bbox(ndvi)
projection(ndvi)
```

### <span style="color:purple">Question 3(c)</span>	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
ndvi.rb <- brick("ndvi.nc")
meanndvi <- mean(ndvi.rb)
plot(meanndvi)
writeRaster(meanndvi,filename = "meanndvi.tif",overwrite=TRUE)
maxndvi <- max(ndvi.rb)
plot(maxndvi)
writeRaster(maxndvi,filename = "maxndvi.tif",overwrite=TRUE)
```

### <span style="color:purple">Question 3(d)</span>			
Plot the maps of monthly NDVI of the year 2001
```{r}
plot(ndvi.rb,13:24)
```

### <span style="color:purple">Question 3(e)</span>	
Create histograms of monthly NDVI of the year 2001.
```{r message=FALSE, warning=FALSE}
hist(ndvi.rb,13:24)
```

### <span style="color:purple">Question 3(f)</span>			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(ndvi.rb,7)
```

```{r}
values <- click(ndvi.rb,n=1, xy=FALSE)
plot(as.vector(values), type="b")
```

