---
title: "Lab 4"
author: "Michael Kezner"
date: October 7, 2017
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <span style="color:red">Question 1</span>
### <span style="color:blue">Part A</span>

**The probability of a team scoring 20 points or fewer:**
```{r 20 or fewer}
pnorm(20, mean = 32, sd = 7)
```
The probability of a team scoring 20 points or fewer is **0.04323813** 

<br>

**Desnity Function:**
```{r 20 or fewer density function}
curve(dnorm(x, 32, 7), xlim = c(11, 53), col = "deepskyblue4", lwd = 2)
cord.x <- c(11, seq(11, 20, 0.01), 20)
cord.y <- c(0, dnorm(seq(11, 20, 0.01), 32, 7), 0)
curve(dnorm(x, 32, 7), xlim = c(11, 53), col = "deepskyblue4", lwd = 2)
polygon(cord.x, cord.y, col = "deepskyblue1")
polygon(cord.x, cord.y, col = 'deepskyblue4', density = 10)
```

### <span style="color:blue">Part B</span>
**The probability of a team scoring more than 35 points:**
```{r more than 35 points}
pnorm(35, 32, 7, lower.tail = FALSE)
```
The probability of a team scoring more than 35 points is **0.3341176**

<br>

**Density Function:**
```{r more than 35 density function}
curve(dnorm(x, 32, 7), xlim = c(11, 53), col = 'darkorchid4', lwd = 2)
cord.x1 <- c(35, seq(35, 53, 0.01), 53)
cord.y1 <- c(0, dnorm(seq(35, 53, 0.01), 32, 7), 0)
curve(dnorm(x, 32, 7), xlim = c(11, 53), col = 'darkorchid4', lwd = 2)
polygon(cord.x1, cord.y1, col = 'orchid1')
polygon(cord.x1, cord.y1, col = 'darkorchid3', density = 10)
```

### <span style="color:blue">Part C</span>
**The probability of a team scoring between 20 and 40 points**
```{r between 20 and 40 points}
diff(pnorm(c(20, 40), 32, 7))
```
The probability of a team scoring between 20 and 40 points is **0.8302129**

<br>

**Density Function:**
```{r between 20 and 40 density function}
curve(dnorm(x, 32, 7), xlim = c(11, 53), col = 'mediumseagreen', lwd = 2)
cord.x2 <- c(20, seq(20, 40, 0.01), 40)
cord.y2 <- c(0, dnorm(seq(20, 40, 0.01), 32, 7), 0)
curve(dnorm(x, 32, 7), xlim = c(11, 53), col = 'mediumseagreen', lwd = 2)
polygon(cord.x2, cord.y2, col = 'mediumspringgreen')
polygon(cord.x2, cord.y2, col = 'mediumseagreen', density = 10)
```

## <span style="color:red">Question 2</span>
### <span style="color:blue">Part A</span>
**Percent of posts that get less than three comments:**
```{r less than three comments}
pexp(3, rate = (1/10))
```
The percent of posts that get less than three comments is **25.918%**

<br>

**Density Function:**
```{r less than three comments density function}
curve(dexp(x, 0.1), to = 100, xlim = c(0, 50), lwd = 2, col = "chocolate4")
xcord <- c(0, seq(0, 3, 0.01), 3)
ycord <- c(0, dexp(seq(0,3, 0.01), 0.1), 0)
curve(dexp(x, 0.1), to = 100, xlim = c(0, 50), lwd = 2, col = "chocolate4")
polygon(xcord, ycord, col = 'chocolate1')
polygon(xcord, ycord, col = 'chocolate4', density = 10)
```

### <span style="color:blue">Part B</span>
**The percent of posts that get more than 20 comments:**
```{r more than 20 comments}
pexp(20, rate = 0.1, lower.tail = FALSE)
```
The percent of posts that get more than 20 comments is **13.434%**

<br>

**Density Function:**
```{r more than 20 density function}
curve(dexp(x, 0.1), to = 60, xlim = c(0, 50), lwd = 2, col = 'palevioletred4')
xcord1 <- c(20, seq(20, 60, 0.01), 50)
ycord1 <- c(0, dexp(seq(20, 60, 0.01), 0.1), 0)
curve(dexp(x, 0.1), to = 60, xlim = c(0, 50), lwd = 2, col = 'palevioletred4')
polygon(xcord1, ycord1, col = 'palevioletred1')
polygon(xcord1, ycord1, col = 'palevioletred4', density = 10)
```

### <span style="color:blue">Part C</span>
**The percent of posts that get between five and ten comments:**
```{r between five and ten comments}
diff(pexp(c(5, 10), 0.1))
```
The percent of posts that get between five and ten comments is **23.865%**

<br>

**Density Function:**
```{r between five and ten density function}
curve(dexp(x, 0.1), to = 60, xlim = c(0, 50), lwd = 2, col = 'turquoise4')
xcord2 <- c(5, seq(5, 10, 0.01), 10)
ycord2 <- c(0, dexp(seq(5, 10, 0.01), 0.1), 0)
curve(dexp(x, 0.1), to = 60, xlim = c(0, 50), lwd = 2, col = 'turquoise4')
polygon(xcord2, ycord2, col = 'turquoise')
polygon(xcord2, ycord2, col = 'turquoise4', density = 10)
```

## <span style="color:red">Question 3</span>
### <span style="color:blue">Part A</span>

**First raster with assigned values based on random values from a uniform distribution:**
```{r first raster uniform distribution}
library(raster)
library(sp)
ras1 <- raster(nrows = 30, ncols = 30, xmn = 0, xmx = 30, ymn = 0, ymx = 30)
ras1
```

<br>

**First Raster Plot:**
```{r first raster plot}
ras1[] <- runif(ncell(ras1))
plot(ras1)
```

**First Raster Histogram:**
```{r first raster histogram}
hist(ras1)
```

### <span style="color:blue">Part B</span>
**Second raster with assigned values based on random values from a normal distribution:**
```{r second raster normal distribution}
ras2 <- raster(nrows = 30, ncols = 30, xmn = 0, xmx = 30, ymn = 0, ymx = 30)
ras2
```

<br>

**Second Raster Plot:**
```{r second raster plot}
ras2[] <- rnorm(ncell(ras2))
plot(ras2)
```
**Second Raster Histogram:**
```{r second raster histogram}
hist(ras2)
```

### <span style="color:blue">Part C</span>
**Third raster with assigned values based on cell-by-cell addition of the first and second raster:**
```{r third raster}
ras3 <- ras1 + ras2
ras3
```

<br>

**Third Raster Plot:**
```{r third raster plot}
plot(ras3)
```
**Third Raster Histogram:**
```{r third raster histogram}
hist(ras3)
```

### <span style="color:blue">Part D</span>
**Mean value of the third raster:**
```{r third raster mean}
cellStats(ras3, "mean")
```
The mean value of the third raster is **0.5121868**

<br>

**Reclassifying the third raster as a binary image:**
```{r third raster reclass}
ras4 <- ras3
ras4[ras3 > 0.5121868] <- 1
ras4[ras3 <= 0.5121868] <- 0
plot(ras4)
```

**Saving the reclassifyed image to my working directory as a TIFF image:**
```{r reclassed .tif image}
library(rgdal)
writeRaster(ras4, filename = "lab4ras4.tif", overwrite = TRUE)
```

