---
title: "Lab 3"
author: "Michael Kezner"
date: September 30, 2017
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <span style="color:red">Question 1</span>
### <span style="color:blue">Part A</span>
```{r Cars93, echo=FALSE}
library(MASS)
df <- Cars93
```
<br>

#### Training Dataset
```{r training, echo=FALSE}
set.seed(1)
training <- df[sample(nrow(df), 40, replace=FALSE), ]
```

```{r training table, echo=FALSE, message=FALSE, warning=FALSE}
library(DT)
datatable(training,options = list(pageLength = 5))
```

**Training Dataset saved as a csv file**
```{r training csv}
write.csv(training, "training.csv")
training2 <- read.csv("training.csv")
```
<br>

#### Test Dataset
```{r test, echo=FALSE}
test <- df[!rownames(df) %in% rownames(training),]
```

```{r test table, echo=FALSE}
datatable(test,options = list(pageLength = 5))
```


**Test Dataset saved as a csv file**
```{r test csv}
write.csv(test, "test.csv")
test2 <- read.csv("test.csv")
```
<br>

### <span style="color:blue">Part B</span>

**Number of USA cars in the Training Dataset**
```{r USA, echo=FALSE}
USA <- training[training$Origin == "USA", ]
nrow(USA)
```

**Number of Non-USA cars in the Training Dataset**
```{r Non-USA, echo=FALSE}
NONUSA <- training[training$Origin == "non-USA", ]
nrow(NONUSA)
```
<br>


## <span style="color:red">Question 2</span>
### <span style="color:blue">Part A</span>

#### Sample 10 births

```{r 10 births}
z <- sample(c("boy", "girl"), 10, replace = TRUE, prob = c(0.515, 0.485))
z
```

**How many boys were there?**
```{r boys10, echo=FALSE}
sum(z=="boy")
```

**How many girls were there?**
```{r girls10, echo=FALSE}
sum(z=="girl")
```
<br>

### <span style="color:blue">Part B</span>
#### Sample 1000 births
```{r 1000 births}
s <- sample(c("boy", "girl"), 1000, replace = TRUE, prob = c(0.515, 0.485))
```

**How many boys were there?**
```{r boys1000, echo=FALSE}
sum(s=="boy")
```

**How many girls were there?**
```{r girls1000, echo=FALSE}
sum(s=="girl")
```
<br>

## <span style="color:red">Question 3</span>
### <span style="color:blue">Part A</span>

**Probability of having exactly 3 boys out of 10 births:**
```{r exactly 3 boys, echo=FALSE}
b <- dbinom(3, size =  10, prob = 0.515)
b
```
<br>

### <span style="color:blue">Part B</span>

**Probability of having 8 or more boys out of 10 births**
```{r 8 or more boys, echo=FALSE}
b2 <- diff(pbinom(c(8, 10), size = 10, prob = 0.515))
b2
```
<br>

### <span style="color:blue">Part C</span>
**Density function plot for boy births out of 10**
```{r density plot, echo=FALSE}
n <- 10
x <- 0:n
y <- dbinom(x, size = n, prob = 0.515)
plot(x,y, type = "b", pch = 16, col = "blue")
```

**Cumulative probability function plot for boy births out of 10**
```{r cumulative plot, echo=FALSE}
n <- 10
x <- 0:10
y1 <- pbinom(x, size = n, prob = 0.515)
plot(x,y1, type = "b", pch = 16, col = "red")
```
<br>

## <span style="color:red">Question 4</span>
### <span style="color:blue">Part A</span>

**Probability that exactly four cars arrive in the next 10 seconds**
```{r four cars, echo=FALSE}
dpois(4, lambda = 5)
```
<br>

### <span style="color:blue">Part B</span>

**Probability that more than five cars arrive in the next 10 seconds**
```{r more than five cars, echo=FALSE}
ppois(4, lambda = 5, lower.tail = FALSE)
```
<br>

### <span style="color:blue">Part C</span>

**Density function plot of 0-20 cars arriving in the next 10 seconds**
```{r car plot, echo=FALSE}
n1 <- 20
x1 <- 0:n1
y2 <- dpois(x1, lambda = 5)
plot(x1,y2, type = "b", pch = 16, col = "purple")
```

